### 2023-10-24 paper/lensbans.tex
Adjoints folds

\newsavebox{\lbansbox}
\newcommand{\lbansmacro}{%
\begin{tikzpicture}[baseline=0.45ex,xscale=0.006em,yscale=0.012ex]%
% curvey bananas
%\draw[solid,join=round,fill=yellow] (2,0) to[out=140,in=-90] (0,3) to[out=90,in=-140] (2,6) -- (2.1,5.9) to[out=-120,in=90] (1.2,3) to[out=-90,in=120] (2.1,0.1) -- cycle;%
% fitted lenses
\draw[solid,join=round] (1.8,6) -- (1.5,5.9) to[out=-120, in=90] (0.7,3) to[out=-90, in=120] (1.5,0.1) -- (1.8,0) -- cycle;%
\end{tikzpicture}}
\savebox{\lbansbox}{\lbansmacro}
\newcommand{\lbans}{\mathopen{\usebox{\lbansbox}\mspace{1mu}}}

\newsavebox{\rbansbox}
\newcommand{\rbansmacro}{%
\begin{tikzpicture}[baseline=0.45ex,xscale=-0.006em,yscale=0.012ex]%
% curvey bananas
%\draw[solid,join=round,fill=yellow] (2,0) to[out=140,in=-90] (0,3) to[out=90,in=-140] (2,6) -- (2.1,5.9) to[out=-120,in=90] (1.2,3) to[out=-90,in=120] (2.1,0.1) -- cycle;%
% fitted lenses
\draw[solid,join=round] (1.8,6) -- (1.5,5.9) to[out=-120, in=90] (0.7,3) to[out=-90, in=120] (1.5,0.1) -- (1.8,0) -- cycle;%
\end{tikzpicture}}
\savebox{\rbansbox}{\rbansmacro}
\newcommand{\rbans}{\mathclose{\mspace{1mu}\usebox{\rbansbox}}}

\newsavebox{\llensbox}
\newcommand{\llensmacro}{%
\begin{tikzpicture}[baseline=0.45ex,xscale=0.006em,yscale=0.012ex]%
\draw[solid,join=round] (1.4,0) -- (0,0) -- (0,6) -- (1.4,6) -- (1.5,5.9) to[out=-120, in=90] (0.7,3) to[out=-90, in=120] (1.5,0.1) -- cycle;%
\end{tikzpicture}}
\savebox{\llensbox}{\llensmacro}
\newcommand{\llens}{\mathopen{\usebox{\llensbox}\mspace{1mu}}}

\newsavebox{\rlensbox}
\newcommand{\rlensmacro}{%
\begin{tikzpicture}[baseline=0.45ex,xscale=-0.006em,yscale=0.012ex]%
\draw[solid,join=round] (1.4,0) -- (0,0) -- (0,6) -- (1.4,6) -- (1.5,5.9) to[out=-120, in=90] (0.7,3) to[out=-90, in=120] (1.5,0.1) -- cycle;%
\end{tikzpicture}}
\savebox{\rlensbox}{\rlensmacro}
\newcommand{\rlens}{\mathclose{\mspace{1mu}\usebox{\rlensbox}}}