### 2024-12-04 ESOP/rebuttal.txt
rebuttal

Reviewer A
------------------------------------
I didn't get a clear sense from the paper of the depth and signiicance of the
contribution.  Setoid based reasoning is well understood (if painful).
Containers have been studied and also formalized before, though I think mostly
in Agda and perhaps mostly in settings assuming axioms that might complicate
code extraction.  The material on hylomorphims and exampls itself is not new.
So it seems what is new is how to put the pieces together to avoid the various
pitfalls in formalizing hylomorphisms in Coq to maintain the ability to generate
OCaml code.  The code that is generated still looks like it would need
additional work to incorporate into a program dealing with say ordinary lists or
trees defined using OCaml datatypes rather than as container equivalents.  So
I'm not fully persuaded that the benefits of the end results make the paper
compelling on their own either.

Reviewer B
------------------------------------
A criticism is that I am struggling to position this work in a particular area.
Is this an elaborate new Coq library that's going to be useful to lots of
people? Does it provide any major take-away lessons for Coq programming? I'm not
sure. Above I have guessed what the take-aways are, but this is not summarized
in the introduction.

Reviewer C
------------------------------------
I could not understand what is the real motivation of this study. Is it a case
study of various Coq features, including the OCaml-code extraction function? Or,
in a real situation, does an OCaml programmer use this to write a program? For
the latter, why does using Coq help with which problem of hylomorphisms?

Although the paper gives gentle explanations here and there, I felt that there
is little self-contained and detailed explanation on the termination issue.

Overall, I felt that the explanation of the motivation is insufficient, while
there seems to be no serious problem in the paper, and the authors at least work
out some amount of a case study on OCaml extraction from Coq. I have no strong
opinion for rejection or acceptance, and I think that there seems to be no
problem if this paper is accepted.