### 2024-12-04 ESOP/rebuttal.txt
minor

We would like to thank the reviewers for their constructive feedback, which has
helped us clarify our contributions and refine the positioning of our work.
Below, we address the concerns raised by each reviewer, emphasizing the novelty,
motivation, and significance of our contributions.

Overall we think the reviewers have a sound understanding of the paper. We would
like to emphasize that our approach uniquely combines a formal and sound OCaml
extraction of recursion schemes within Coq, addressing termination challenges
without relying on axioms. This is the first formalisation of recursion schemes
within a type theory which ensures termination of definitions as opposed to
current presentations of recursion schemes in, e.g., Haskell which can be broken
by using non-terminating arguments to the recursion scheme.

Using Coq extraction mechanism we can provide the functional programmer with a
termination guarantee and at the same time provide the Coq programmer with a
library which accepts more recursive definitions.

Revisions will highlight practical insights for Coq programming, expand on
termination handling, and articulate the broader significance of bridging formal
reasoning with practical code generation.

We will refine the introduction to clearly summarize these
contributions and position the work within the broader landscape.

Summary of Planned Revisions
---------------------------------------------------------------------------
In response to the reviewers' comments, we will make the following changes:

- Clarify Motivation and Contributions: Revise the introduction to clearly
articulate the motivation, contributions, and broader significance of the work,
positioning it as both a methodological advancement and a practical framework.

- Highlight Novelty: Emphasize how our work goes beyond prior formalizations by
addressing specific challenges in Coq related to termination and extraction,
which have not been explored in depth in Agda or other systems.

- Explanation of Termination: Include a detailed and self-contained discussion of
termination issues and how our framework resolves them.  Future Work: Add a
discussion on extending the framework for better integration with standard OCaml
datatypes and libraries.

We hope these clarifications address the reviewers' concerns and demonstrate the
value of our contributions. Thank you for your thoughtful and detailed feedback.

Reviewer A
------------------------------------
Comment:
> Setoid-based reasoning is well understood (if painful). Containers have been
> studied and also formalized before, though I think mostly in Agda...The material
> on hylomorphisms and examples itself is not new...So it seems what is new is how
> to put the pieces together to avoid the various pitfalls in formalizing
> hylomorphisms in Coq to maintain the ability to generate OCaml code.

We agree with the reviewer that setoid reasoning, hylomorphisms and containers
are established topics. As correctly pointed out by the reviewer however they
have never been pieced together. In particular, the whole area of recursion
schemes has only been considered within the broader realm of functional
programming. In this setting, it is still possible to introduce non-termination,
e.g. by feeding the recursion scheme with invalid non-terminating algebras.
Hence recursion scheme have mostly been used to structure programs for reasoning
rather than ensuring formally that they define terminating functions.

This work can be viewed as a first step towards considering the use of recursion
 schemes to help proof assistants to accept more recursive definitions.

The issue of extracting runnable code is perhaps tangential but still helpful
from a programming point of view as 1) OCaml does not have full access to
recursion schemes since there is no way of declaring a functor as easily as for
example in Haskell and 2) it provides

Comment: "The code that is generated still looks like it would need additional
work to incorporate into a program dealing with, say, ordinary lists or trees
defined using OCaml datatypes rather than as container equivalents."

[DAVID TO DO]

Reviewer B
------------------------------------
Comment:
> I am struggling to position this work in a particular area...Is this
> an elaborate new Coq library that’s going to be useful to lots of people? Does
> it provide any major take-away lessons for Coq programming?

Our primary goal is to provide a framework that advances the formalization and
practical usability of recursive schemes like hylomorphisms in Coq, particularly
with a focus on sound OCaml code extraction. While this work includes the
implementation of a Coq library, its broader significance lies in showcasing how
termination-sensitive constructs like hylomorphisms can be handled rigorously
within Coq's framework, offering insights into programming techniques that
balance formal reasoning with practical usability.

Comment:
> Above I have guessed what the take-aways are, but this is not
> summarized in the introduction.

Thank you for highlighting this. We will revise the introduction accordingly.

Reviewer C
------------------------------------
Comment:
> I could not understand what is the real motivation of this study. Is
> it a case study of various Coq features, including the OCaml-code extraction
> function?

Comment:
> In a real situation, does an OCaml programmer use this to write a
> program? For the latter, why does using Coq help with which problem of
> hylomorphisms?

Comment:
> I felt that there is little self-contained and detailed explanation on
> the termination issue.