### 2024-12-04 ESOP/dcastrop_rebuttal.txt
Rebuttal Version Dcastrop

We would like to thank the reviewers for their constructive feedback, which has
helped us clarify our contributions and refine the positioning of our work.

A common theme among the reviewers is the need for a clearer presentation of the
motivation behind our work, which we address here in summary.

The main motivation of our work is to build a verified framework for program
calculation and code extraction. The state-of-the art does not address
divide-and-conquer algorithms, and does not address the extraction of idiomatic
code.

While the individual components of this work are well-known, as correctly noted
by Reviewer A, their integration into a cohesive framework has not been
previously studied. Moreover, their use to formalise a divide-and-conquer
recursion scheme (hylomorphisms) in a type theory has also never been explored.
One key benefit of our approach is that we allow the use of regular Coq proofs
to do program calculation, i.e. derive efficient implementations from simple
specifications. Unlike previous work, we can encode divide-and-conquer
computations, and our mechanisation extracts idiomatic OCaml code. We note in
Section 6.1 that previous formalisations of recursion schemes in proof
assistants focus on maps and folds. Only Abreu et al. (2023) encode divide and
conquer computations, but in a way that does not lead to idiomatic code
extraction (Section 6.2). Code extraction is not straightforward, and it
requires the careful selection of the representations of containers, functors,
etc, such as the use of decidable predicates to define indexed families of
types.

Furthermore, the use of hylomorphisms in a type theory has some benefits
regarding termination proofs. Unlike common approaches to encode non-structural
recursion in a proof assistant, the use of hylomorphisms allows to focus the
proofs on showing that the coalgebra is recursive. This means that termination
proofs become more modular, and can be plugged in to build different
divide-and-conquer computations. Our framework allows programmers to build
libraries of recursive coalgebras for different datatypes, which can be reused
when building other non-structural recursive functions. The encoding of
recursive coalgebras to prove termination has not been explored before.

Our work allows to encode program calculation proofs in a natural way that,
until today, are not machine-checked. Furthermore, we believe our work is a
first step in try to understand how recursion schemes can be integrated in type
theories to tackle termination challenges. We intend to have a more theoretical
follow up addressing corecursive algorithms.

Review A
===========================================================================

> This paper investigates the challenges of formalizing reasoning about
> recursion schemes (cata, ana, hylomorphisms) in Coq in a way that is
> compatible with code extraction.  Other approaches have run into problems,
> first because [...]

It is important to note that, as we point out in Section 6.1, other approaches
do not formalise divide-and-conquer computations. We allow them due to our
encoding of recursive coalgebras.

> The code that is generated still looks like it would need additional work to
> incorporate into a program dealing with say ordinary lists or trees defined
> using OCaml datatypes rather than as container equivalents.

The use of ordinary lists and trees is not only possible, but easy. In fact,
containers are only used internally in our examples, the input/output types use
regular OCaml datatypes. We do that by defining them as initial algebras of the
respective container extension (e.g. ListF, or TreeF in Section 5). By doing
that, we get (for free) that OCaml lists/trees/etc are isomorphic to `Fix
ListF` or `Fix TreeF`.  To further simplify interacting with these container
extensions, it should be straightforward (even if tedious) to automate the
translation between a strictly positive datatype and its container equivalent
(e.g. Section 6 of Abott et al. (2005), or our own encoding of polynomial
functors as containers in Coq).

> p12.  I the definition of comp, I was surprised that the arrows between (B ~>
> C) and (A ~> B) and (A ~> C) are not ordinary function arrows.  I see this is
> also the way it is in the formalization, is there a particular reason it has
> to be this way or just for convenience?

It simplifies the proofs. In particular, if we have that `f =e g`, then we can
trivially conclude that `f \o h =e g \o h`. Otherwise, we would need to do this
proof manually every time it is required.

Review B
===========================================================================

> * "least" fixed point, what does "least" mean here? and anyway that is not what Lambek's lemma says.

[TODO MARCO]

> * "pronounced catamorphism" I think this is functional programming terminology, not category theory

[TODO MARCO]

> * "Note that since Haskell". I didn't understand the point of this paragraph. I do think Haskell is a bad language for demonstrating this.

> Section 2.3. Again, the terminal object might not exist, "greatest" doesn't make sense, etc..

[TODO MARCO: does "greatest" not make sense?? I thought it was standard terminology]

> Page 8. "TreeF". This is not a tree, and it's never used as a tree. It's just a pair of lists with a midpoint.

The least/greatest fixpoint of TreeF is an inductive/coinductive tree.

> Page 9. Quicksort is not a solution to this equation, rather, "sort" is a
> solution. These are functions between sets, and so cannot distinguish
> intensional algorithmic specifics.

The solution to this equation behaves exactly as the naive implementation of
Quicksort in Haskell/OCaml.

I did not understand the recursive anamorphisms paragraph, but it seems to be a digression.

> Does the extra variable x0 have a performance cost?

The extra `x0` can be trivially inlined in a compiler pass. We are also
considering extending Coq's code extraction with manual inlining hints.

Review C
===========================================================================

> I could not understand what is the real motivation of this study.
> does an OCaml programmer use this to write a program?

Our end goal is to build a framework for program verification based on program
calculation techniques. In that sense, our framework is meant to make the use
of program calculation more accessible to programmers. Prior to this work,
using program calculation for divide-and-conquer computations would require
pen-and-paper proofs and a manual translation to Haskell/OCaml code.

> For the latter, why does using Coq help with which problem of hylomorphisms?

Hylomorphisms are not guaranteed to exist, or to be unique, due to termination
issues. If a hylomorphism can be encoded in our framework in Coq, it must exist
and be unique.