### 2024-12-05 ESOP/dcastrop_rebuttal.txt
Including Mike's comments'

We would like to thank the reviewers for their constructive feedback, which has
helped us clarify our contributions and refine the positioning of our work.

A common theme among the reviewers is the need for a clearer presentation of the
motivation behind our work, which we address here in summary.

The main motivation of our work is to build a verified framework for program
calculation and code extraction. The state-of-the art does not address
divide-and-conquer algorithms, and does not address the extraction of idiomatic
code.

While the individual components of this work are well-known, as correctly noted
by Reviewer A, their integration into a cohesive framework has not been
previously studied. Moreover, their use to formalise a divide-and-conquer
recursion scheme (hylomorphisms) in a type theory has also never been explored.
One key benefit of our approach is that we allow the use of regular Coq proofs
to do program calculation, i.e. derive efficient implementations from simple
specifications. Unlike previous work, we can encode divide-and-conquer
computations, and our mechanisation extracts idiomatic OCaml code. We note in
Section 6.1 that previous formalisations of recursion schemes in proof
assistants focus on maps and folds. Only Abreu et al. (2023) encode divide and
conquer computations, but in a way that does not lead to idiomatic code
extraction (Section 6.2). Code extraction is not straightforward, and it
requires the careful selection of the representations of containers, functors,
etc, such as the use of decidable predicates to define indexed families of
types.

Furthermore, the use of hylomorphisms in a type theory has some benefits
regarding termination proofs. Unlike common approaches to encode non-structural
recursion in a proof assistant, the use of hylomorphisms allows to focus the
proofs on showing that the coalgebra is recursive. This means that termination
proofs become more modular, and can be plugged in to build different
divide-and-conquer computations. Our framework allows programmers to build
libraries of recursive coalgebras for different datatypes, which can be reused
when building other non-structural recursive functions. The encoding of
recursive coalgebras to prove termination has not been explored before.

Our work allows to encode program calculation proofs in a natural way that,
until today, are not machine-checked. Furthermore, we believe our work is a
first step in try to understand how recursion schemes can be integrated in type
theories to tackle termination challenges. We intend to have a more theoretical
follow up addressing corecursive algorithms.

Review A
===========================================================================

> This paper investigates the challenges of formalizing reasoning about
> recursion schemes (cata, ana, hylomorphisms) in Coq in a way that is
> compatible with code extraction.  Other approaches have run into problems,
> first because [...]

It is important to note that, as we point out in Section 6.1, other approaches
do not formalise divide-and-conquer computations. We allow them due to our
encoding of recursive coalgebras.

> The code that is generated still looks like it would need additional work to
> incorporate into a program dealing with say ordinary lists or trees defined
> using OCaml datatypes rather than as container equivalents.

@@MIKE: This is a valid observation, in that the use of containers in our examples
results in a slight mismatch between the extracted OCaml code and, for example,
existing code/libraries that are expecting more ordinary datatypes for trees or
lists. This could be ameliorated with either a minor amount of marshalling, or
even a straightforward automatic translation. However, the fact that this
extracted code contains no unsafe coercions, and contains minimal noise imposed
by the details of the Coq formalism, make it possible at all for the resulting
OCaml code to be usable. As we detail in the paper, prior related work does not
meet those criteria.

@@David: The use of ordinary lists and trees is not only possible, but easy. In fact,
containers are only used internally in our examples, the input/output types use
regular OCaml datatypes. We do that by defining them as initial algebras of the
respective container extension (e.g. ListF, or TreeF in Section 5). By doing
that, we get (for free) that OCaml lists/trees/etc are isomorphic to `Fix
ListF` or `Fix TreeF`.  To further simplify interacting with these container
extensions, it should be straightforward (even if tedious) to automate the
translation between a strictly positive datatype and its container equivalent
(e.g. Section 6 of Abott et al. (2005), or our own encoding of polynomial
functors as containers in Coq).

> p12.  I the definition of comp, I was surprised that the arrows between (B ~>
> C) and (A ~> B) and (A ~> C) are not ordinary function arrows.  I see this is
> also the way it is in the formalization, is there a particular reason it has
> to be this way or just for convenience?

It simplifies the proofs. In particular, if we have that `f =e g`, then we can
trivially conclude that `f \o h =e g \o h`. Otherwise, we would need to do this
proof manually every time it is required.

Review B
===========================================================================

> * "least" fixed point, what does "least" mean here? and anyway that is not what Lambek's lemma says.

We don't quite understand this comment.

We define least fixed-point for F as the initial object in the category F-iso of
fixed-points for F and F-iso homomorphisms.  By Lambek's lemma an initial
F-algebra is a fixed-point, but in particular also the least fixed-point,
because we can take any iso X ~ FX, and prove that there exists a unique F-iso
morphism (mu F ---> X) simply because X is an F-algebra.

An initial F-algebra is also a fixed-point for F and in particular is also the least.

> * "pronounced catamorphism" I think this is functional programming terminology, not category theory

we agree, although despite the FP community has adopted their own terminology
what they're indeed referring to the exact same thing.

> * "Note that since Haskell". I didn't understand the point of this paragraph.
    I do think Haskell is a bad language for demonstrating this.

The point here is that Haskell has equi-recursive types which are both inductive
and coinductive data types.  However, when programming with recursion schemes it
is common to make inductive/coinductive types distinguished by defining a Fix
and CoFix operator and then imposing that catamorphisms are used on Fix and
anamorphisms are used on CoFix.

> Section 2.3. Again, the terminal object might not exist, "greatest" doesn't make sense, etc..

The argument might just be the dual of what we explained earlier.
The terminal F-coalgebra might not exist, but if it does it corresponds to the greatest fixed-point of the functor F.

> Page 8. "TreeF". This is not a tree, and it's never used as a tree. It's just a pair of lists with a midpoint.

The least/greatest fixpoint of TreeF is an inductive/coinductive tree. The name TreeF it refers
to the fact that it is the functor giving rise to the a tree via the fixed-point operator.

> Page 9. Quicksort is not a solution to this equation, rather, "sort" is a
> solution. These are functions between sets, and so cannot distinguish
> intensional algorithmic specifics.

The solution to this equation behaves exactly as the naive implementation of
Quicksort in Haskell/OCaml.

> I did not understand the recursive anamorphisms paragraph, but it seems to be a digression.

We do use recursive anamorphisms in page 18, paragraph "On the subtype of finite
elements".  In particular an anamorphism applied to a recursive coalgebra yields
a finite F-tree (for a functor F).

Consider the Coq type

Inductive FinF : GFix F -> Prop

Then {x : GFix F | FinF x} is the subtype of finite elements (of a coinductive
data type) which is isomorphic to its inductive counterpart.

> Does the extra variable x0 have a performance cost?

The extra `x0` can be trivially inlined in a compiler pass. We are also
considering extending Coq's code extraction with manual inlining hints.

Review C
===========================================================================
We appreciate Reviewer C's thoughtful question regarding the motivation of our
study. To clarify, the central contribution of our work lies in proposing a
cohesive framework for recursion schemes in a proof assistant such as Coq, and
the primary *application* of this framework is the verification of
divide-and-conquer algorithms in Coq with clean extraction to readable and
idiomatic OCaml. In contrast, prior work on recursion schemes in Haskell has
largely relied on informal, pen-and-paper proofs and manual, unverified
implementations in Haskell, leaving a gap in the ability to produce formally
verified and automatically extracted implementations. Our framework addresses
this gap by integrating formal verification and code generation into a single,
cohesive process.

> I could not understand what is the real motivation of this study.
> does an OCaml programmer use this to write a program?

Our end goal is to build a framework for program verification based on program
calculation techniques. In that sense, our framework is meant to make the use
of program calculation more accessible to programmers. Prior to this work,
using program calculation for divide-and-conquer computations would require
pen-and-paper proofs and a manual translation to Haskell/OCaml code.

> For the latter, why does using Coq help with which problem of hylomorphisms?

Hylomorphisms are not guaranteed to exist, or to be unique, due to termination
issues. If a hylomorphism can be encoded in our framework in Coq, it must exist
and be unique.